---
title: '[BRIGHTEN Study](http://brightenstudy.com/study-description.html) Mobility
  Data Analysis'
author: "Abhishek Pratap @ UW / Sage Bionetworks"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
rm(list=ls())
library("install.load")
install_load("data.table", "gdata", "ggplot2", "e1071", "grid", "reshape2", "randomForest")
install_load("plyr", "tidyverse", "ROCR", "caret", "doMC", "scales", "sjPlot", "caret", "gridExtra")
install_load("ranger", "caret", "printr", "ggthemes" ,"synapseClient", "ROCR", "doMC", "ggpubr")

registerDoMC(4)
synapseLogin()
```

```{r, echo=F}
knitr::opts_chunk$set(echo = F,  message=FALSE, warning=FALSE)
knitr::opts_chunk$set(comment = NA, results = "asis", comment = NA, tidy = F)
```

```{r}
source("../loadData.R")
ls()
phq2 <- phq2  %>% filter(!is.na(sum_phq2))
phq9 <- phq9  %>% filter(!is.na(sum_phq9))

n_distinct(phq9$brightenid)
n_distinct(phq2$brightenid)
n_distinct(passive_data$brightenid)
n_distinct(c(phq2$brightenid, passive_data$brightenid))
```



```{r}
#select only the mobility 
mobilityfeatures <- passive_data %>% 
  select(brightenid, day, week, mobility, mobility_radius, passive_date_pacific) %>%
  filter(!(is.na(mobility) | is.na(mobility_radius))) %>%
  mutate(mobRadius_trans = log10(mobility_radius+.0001),
         mobility_trans = log10(mobility+.0001),
         passive_date_pacific = as.character(passive_date_pacific))

#merge with the study meta data
tmp_metaData <- metaData %>% select(Gender, Age, brightenid ) %>% 
  filter(!is.na(brightenid),
         !is.na(Gender),
         Gender != '')
mobilityfeatures <- merge(tmp_metaData, mobilityfeatures, by="brightenid")
mobilityfeatures <- mobilityfeatures %>% filter(! (is.na(Age) | Age == '')) %>%
  filter(Age >=18)
n_distinct(mobilityfeatures$brightenid)
```

### After cleaning number of individuals for whom we have mobility data
```{r}
n_distinct(mobilityfeatures$brightenid)
```


```{r}
## Compare complaince for mobility and PHQ2 and PHQ9
compliance_mobilityfeatures <- mobilityfeatures %>% mutate(brightenid = as.character(brightenid)) %>%
  group_by(week, brightenid) %>% 
  summarise(n = length(unique(day)),
            survey = 'mobility')

compliance_phq2 <- phq2 %>% mutate(brightenid = as.character(brightenid)) %>% 
  group_by(week, brightenid) %>% 
  summarise(n = length(unique(day)),
            survey = 'phq2')
```


```{r , eval=F}
### Merged Passive and PHQ2 data
tmp_phq2 <- phq2 %>%  mutate(start = as.character(start), day = day-1) %>% select(-week)
mobilityfeatures <- mobilityfeatures %>% dplyr::mutate(brightenid = as.character(brightenid))
intersect(colnames(mobilityfeatures), colnames(tmp_phq2))
mobilityfeatures_n_phq2 <- merge(mobilityfeatures, tmp_phq2)
```

### After cleaning number of individuals for whom we have mobility and PHQ2 data
```{r}
n_distinct(mobilityfeatures_n_phq2$brightenid)
```


### Filter to users with enough data
```{r}
dataPerUser <- mobilityfeatures_n_phq2 %>% group_by(brightenid) %>% 
  summarise(n = length(unique(day)),
            survey = 'passive')

hist(dataPerUser$n)
#select individuals with atleast 15 data points 

SELECTED_USERS <- dataPerUser %>% filter( n >= 14) %>% .$brightenid %>% unique()
mobilityfeatures_n_phq2_flt <- mobilityfeatures_n_phq2 %>% filter(brightenid %in% SELECTED_USERS)
```

```{r}
n_distinct(mobilityfeatures_n_phq2_flt$brightenid)
```


### Overall mobility var
```{r}
summary(mobilityfeatures_n_phq2$mobility)
sd(mobilityfeatures_n_phq2$mobility)

summary(mobilityfeatures_n_phq2$mobility_radius)
sd(mobilityfeatures_n_phq2$mobility_radius)

summary(mobilityfeatures_n_phq2$sum_phq2)
sd(mobilityfeatures_n_phq2$sum_phq2)
```




----------------

In BRIGHTEN Data we have mainly two passive features for mobility as generarted by ginger.io (the app used to capture passive data)

1. Mobility 
2. Mobility Radius

We see significant correlation betweeen these two vars

```{r fig.height=5, fig.width=5, fig.align='center'}
library(ggpubr)
ggscatter(data=mobilityfeatures_n_phq2_flt, x="mobRadius_trans", y="mobility_trans",
          size = 1,
          add = "reg.line",  # Add regressin line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          cor.coef = TRUE # Add correlation coefficient
          )
```

----------


Across the full cohort we **don't** see a strong trend. However we need to look at individual N-of-1 level to see if there some interesting trends for some people

```{r fig.height=5, fig.width=10, fig.align='center'}
p1 <- ggscatter(data=mobilityfeatures_n_phq2_flt, x="mobRadius_trans", y="Age",
          size = 1,
          add = "reg.line",  # Add regressin line
          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
          conf.int = TRUE, # Add confidence interval
          cor.coef = TRUE # Add correlation coefficient
          )

p2 <- ggscatter(data=mobilityfeatures_n_phq2_flt, x="mobRadius_trans",
          y="sum_phq2",
          size = 1,
          add = "reg.line",  # Add regressin line
          add.params = list(color = "blue", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE
          )
gridExtra::grid.arrange(p1,p2,nrow=1)
```


#### Difference between mobility across gender 

```{r fig.height=3, fig.width=4, fig.align='center'}
ggplot(data=mobilityfeatures_n_phq2_flt, aes(x=mobRadius_trans, fill=Gender)) + geom_density(alpha=.4)
```


#### Difference between mobility faceted by age group

```{r fig.height=3, fig.width=4, fig.align='center'}
mobilityfeatures_n_phq2_flt$age_group <- cut(mobilityfeatures_n_phq2_flt$Age, c(0,30,40,50,60,70,80), right=T)
#ggplot(data=mobilityfeatures_n_phq2_flt, aes(x=mobRadius_trans, fill=age_group)) + geom_density(alpha=.4)
ggplot(data=mobilityfeatures_n_phq2_flt, aes(x=mobility_trans, fill=age_group)) + geom_density(alpha=.4)
```


--------------------


### N-of-1 Analysis

#### 1. Correlation between Mobility Radius AND SUM_PHQ2 for an individual

Top 12 users with significant correlation with mobility radius. There is individual level heterogenity with some users having +ve correlation to mobility radius with increasing depression severity as measured by PHQ2. 



```{r}
tmp_selected_users <- unique(mobilityfeatures_n_phq2_flt$brightenid)

# cor_mobRadiusTrans <- ldply(tmp_selected_users, function(userId){
#   n1 <- mobilityfeatures_n_phq2_flt %>% filter(brightenid == userId)
#   tmp_cor <- cor(n1$mobRadius_trans, n1$sum_phq2, method="spearman")
#   tmp_pval <- cor.test(n1$mobRadius_trans, n1$sum_phq2, method="spearman")['p.value']
#   data.frame(cor=tmp_cor, pval=tmp_pval, brightenid = userId)
# })
#cor_mobRadiusTrans['pval.adj'] <- p.adjust(cor_mobRadiusTrans$p.value, method="fdr")
#sum(cor_mobRadiusTrans$p.value < .05, na.rm = T)

cor_MobilityTrans <- ldply(tmp_selected_users, function(userId){
  n1 <- mobilityfeatures_n_phq2_flt %>% filter(brightenid == userId)
  tmp_cor <- cor(n1$mobility_trans, n1$sum_phq2, method="spearman")
  tmp_pval <- cor.test(n1$mobility_trans, n1$sum_phq2,  method="spearman")['p.value']
  data.frame(cor=tmp_cor, pval=tmp_pval, brightenid = userId)
})

cor_MobilityTrans['pval.adj'] <- p.adjust(cor_MobilityTrans$p.value, method="fdr")
sum(cor_MobilityTrans$p.value < .05, na.rm = T)
sum(cor_MobilityTrans$pval.adj < .05, na.rm = T)

tmp <- cor_MobilityTrans %>% filter(p.value < .05) %>% arrange(p.value, desc(abs(cor)))
tmp_selected_users <- as.character(tmp$brightenid[c(1,4,3,2)])


tmp_plots <- lapply(1:length(tmp_selected_users), function(index){
  userId <- tmp_selected_users[index]
  n1 <- mobilityfeatures_n_phq2_flt %>% dplyr::filter(brightenid == tmp_selected_users[index])
  ggscatter(data=n1, y="mobility_trans",
          x="sum_phq2",
          size = .3,
          add = "reg.line",  # Add regressin line
          add.params = list(color = "grey10", fill = "lightgray"),
          conf.int = TRUE, 
          cor.coef = TRUE,
          cor.method = "spearman",
          cor.coef.size = 2.5,
          cor.coeff.args = list(method = "pearson", fontface="italic",
                                label.x.npc = "left", colour = "black",
                                label.y.npc = "bottom")
          ) + ggtitle(paste0(letters[index], '.')) + ylab('Mobility (log10)') + xlab('PHQ-2 score') + theme(text = element_text(size=8), plot.title=element_text(hjust=0, face="bold"))
})
pl <- gridExtra::arrangeGrob(grobs=tmp_plots, ncol=2)
ggsave("plots/mobility.png", pl, height=5, width=7, units="in")
ggsave("plots/mobility.tiff", pl, height=5, width=7, units="in")
```




```{r}
##### At Cohort level 
library(ranger)
library(caret)
colnames(mobilityfeatures_n_phq2_flt)
tmp_data <- mobilityfeatures_n_phq2_flt %>% select( -day,-week, -Age, -start, -mobRadius_trans, -mobility_radius) 
colnames(tmp_data)
tmp_data <- tmp_data[complete.cases(tmp_data),]

set.seed(18989)
library(doMC)
registerDoMC(4)
res <- ldply(seq(1:50), .parallel = T, .fun = function(x) {
  users <- unique(tmp_data$brightenid)
  train_users <- sample(users, .70*length(users))
  train <- tmp_data %>% filter(brightenid %in% train_users)
  test <- tmp_data %>% filter(! (brightenid %in% train_users))
  mod.fit <- ranger(sum_phq2 ~ . , data=train, importance="permutation")
  preds <- predict(mod.fit, test)
  testRMSE <- sqrt(mean((test$sum_phq2 - preds$predictions)^2))
  total_variation <- sum((test$sum_phq2 - mean(test$sum_phq2))^2)
  reg_error <- sum((test$sum_phq2 - preds$predictions)^2)
  testRsq <- 1 - (reg_error / total_variation)
  data.frame(testRMSE=testRMSE, testRsq=testRsq)
})

median(res$testRsq)
mean(res$testRsq)
sd(res$testRsq)

```

-----------------


#### 2. Correlation between Mobility Radius AND SUM_PHQ9 for an individual
The correlation/trend with PHQ9 is much more prominent than PHQ2. Note some individuals show positive correlation but their PHQ9 values are significantly lower.


```{r fig.height=10, fig.width=12}
tmp_phq9 <-  as.data.frame(phq9) %>% select(brightenid, sum_phq9, week)
intersect(colnames(tmp_phq9), colnames(mobilityfeatures))
str(tmp_phq9)
str(mobilityfeatures)
dim(tmp_phq9)
dim(mobilityfeatures)
tmp_mobF <- merge(mobilityfeatures,tmp_phq9)
dim(tmp_mobF)

selected_users <- tmp_mobF %>% group_by(brightenid) %>% summarise(numDataPoints = n()) %>% 
  filter(numDataPoints >= 15) %>% .$brightenid
tmp_mobF <- tmp_mobF %>% filter(brightenid %in% selected_users)

cor_mobbility_phq9 <- ldply(selected_users, function(userId){
  n1 <- tmp_mobF %>% filter(brightenid == userId)
  tmp_cor <- cor(n1$mobility_trans, n1$sum_phq9, method="spearman")
  tmp_pval <- cor.test(n1$mobRadius_trans, n1$sum_phq9, method="spearman")['p.value']
  data.frame(cor=tmp_cor, pval=tmp_pval, user_id = userId)
})

View(cor_mobbility_phq9)
sum(cor_mobbility_phq9$p.value < .05, na.rm = T)
cor_mobbility_phq9['pval.adj'] <- p.adjust(cor_mobbility_phq9$p.value, method="fdr")
sum(cor_mobbility_phq9$pval.adj < .05, na.rm = T)

# phq9_res <- phq9_res %>% filter(p.value < .05) %>% arrange(p.value)
# phq9_res <- phq9_res[1:6,]
# 
# tmp_plots <- lapply(phq9_res$user_id, function(userId){
#   n1 <- tmp_mobF %>% filter(user_id == userId)
#   ggscatter(data=n1, y="MobRadius_trans",
#           x="sum_phq9",
#           size = 1,
#           add = "reg.line",  # Add regressin line
#           add.params = list(color = "blue", fill = "lightgray"),
#           conf.int = TRUE, 
#           cor.coef = TRUE
#           ) + ggtitle(userId)
# })
# gridExtra::grid.arrange(grobs=tmp_plots)
```


---------------------------------


#### Next - we try to see if there is a trend between Age, PHQ9, PHQ2 and Mobility. This done through correlation of correlation plot. 

#### A. Correlation between PHQ9/2 and Mobility compared to Age

```{r fig.height=5, fig.width=8}

phq9_correlation_by_age <- tmp_mobF %>% group_by(Age) %>% summarise(cor = cor(sum_phq9, mobility_trans, method="spearman"))

p1 <- ggplot(data=phq9_correlation_by_age, aes(x=Age, y=cor)) + geom_point() + geom_smooth() + ylab('correlation between (Mobility & PHQ9)') + theme_bw()
p1


phq2_correlation_by_age <- mobilityfeatures_n_phq2_flt %>% group_by(Age) %>% summarise(cor = cor(sum_phq2, mobility_trans, method="spearman"))
p2 <- ggplot(data=phq2_correlation_by_age, aes(x=Age, y=cor)) + geom_point() + geom_smooth() + ylab('correlation between (Mobility & PHQ2)') + theme_bw()
gridExtra::grid.arrange(p1,p2, nrow=1)
```




#### Heatmap comparing Median mobility radius / week for each partcipant
```{r fig.height=8, fig.width=10}
#Mobility Radius
x <- mobilityfeatures_n_phq2_flt %>% group_by(brightenid, week) %>%
  summarise(Mobility = log10(median(mobility) + .0001))

x1 <- x %>% tidyr::spread(week, Mobility)
x1 <- x1[complete.cases(x1),]
rownames(x1) <- x1$brightenid
x1$brightenid <- NULL
rowAnnot <- metaData %>% filter(brightenid %in% rownames(x1)) %>%
  select(Age,Gender, brightenid)
rownames(rowAnnot) <- rowAnnot$brightenid
rowAnnot$brightenid <- NULL
pheatmap::pheatmap(t(scale(t(x1))), cluster_cols = F,
                   annotation_row = rowAnnot, show_rownames = F)
```


#### Heatmap comparing Median mobility radius / week for each partcipant
```{r fig.height=8, fig.width=10}
#Mobility Radius
x <- mobilityfeatures %>% group_by(user_id, week_in_study) %>%
  summarise(Mobility = median(Mobility_Radius))
x1 <- tidyr::spread(x, week_in_study, Mobility)
x1 <- x1[complete.cases(x1),]
rownames(x1) <- x1$user_id
x1$user_id <- NULL
rowAnnot <- mdata %>% filter(user_id %in% rownames(x1)) %>%
  select(Age, PHQScore,Gender, user_id)
rownames(rowAnnot) <- rowAnnot$user_id
rowAnnot$user_id <- NULL
pheatmap::pheatmap(t(scale(t(x1))), cluster_cols = F,
                   annotation_row = rowAnnot, show_rownames = F)
```




### N-of-1 Plots

```{r}
#N-of-1
mobilityVar <- mobilityfeatures_n_phq2_flt %>% group_by(user_id) %>% 
  summarise(mob_variance = var(mobility_radius),
            n = length(unique(day))) %>%
  filter(n >= 50)
View(mobilityVar)
  
selected_individuals <- c("11091", "10617" , "11529", "68874")

n_1_plots <- function(participant){
   p1 <- mobilityfeatures_n_phq2_flt %>% filter(user_id == participant)
  tmp_p1 <- ggplot(data=p1, aes(x=day, y=mobRadius_trans)) + geom_line(color="gray") + geom_point(size=.7, color="red") + theme_minimal() + theme(axis.title.x = element_blank(), axis.text.x = element_blank()) + ylab('Mobility radius') 

  tmp_p2 <- ggplot(data=p1, aes(x=day, y=sum_phq2)) + geom_line(color="gray") + geom_point(size=.7, color="blue") + theme_minimal() +  ylab('Daily mood') + xlab('day in study') + scale_x_continuous(limits = c(1,84))
 
  grid.arrange(tmp_p1, tmp_p2, ncol = 1, heights = c(1, 1), top=paste('Participant - ', participant))
}
 
tmp_plots <- lapply(selected_individuals, n_1_plots)
tmp_plot <- grid.arrange(grobs=tmp_plots)
tmp_p
ggsave("n_of_1_plots.png", tmp_plot, height=6, width=8, units="in", dpi=200 )
```





```{r}
# users <- unique(mobilityfeatures_n_phq2_flt$user_id)
# train_users <- sample(users, .60*length(users))
# train <- mobilityfeatures %>% filter(user_id %in% train_users)
# test <- mobilityfeatures %>% filter(! (user_id %in% train_users))
# 
# 
# library(xgboost)
# dtrain <- xgb.DMatrix(data = as.matrix(train %>% select(mobRadius_trans, mobility_trans, Age)), 
#                       label = train$sum_phq2)
# dtest <- xgb.DMatrix(data = as.matrix(test %>% select(MobRadius_trans, Mobility_trans, Age)), 
#                       label = test$sum_phq2)
# 
# 
# 
# bstDense <- xgboost(data = dtrain, max.depth = 3, verbose=2,
#                     eta = 1, nthread = 4, nround = 10, objective= "reg:linear")
# 
# preds <- predict(bstDense, dtest)
# testRMSE <- sqrt(mean((test$sum_phq2 - preds)^2))
# testRMSE
# total_variation <- sum((test$sum_phq2 - mean(test$sum_phq2))^2)
# reg_error <- sum((test$sum_phq2 - preds)^2)
# testRsq <- 1 - (reg_error / total_variation)
# data.frame(testRMSE=testRMSE, testRsq=testRsq)
# 
# dtrain <- xgb.DMatrix(data = as.matrix(train %>% select(-sum_phq2)), label = train$sum_phq2)
```























